% !TEX TS-program = xelatex
% !TEX options = -shell-escape -synctex=1 -interaction=nonstopmode -file-line-error "%DOC%"
\documentclass[twoside,11pt]{report}

\input{preamble}

\setlength{\parskip}{1em}

\allowdisplaybreaks

\begin{document}
% Cover
\onehalfspacing
% \definecolor{light-gray}{gray}{0.7}
% \pagecolor{light-gray}
% \thispagestyle{empty}
% \centerline{\Huge{BYRNE'S EUCLID}}
% \vspace{4ex}
% \centerline{\Large{The First Six Books Of The Elements Of Euclid}}

% \hfill

% \begin{center}
%   \begin{minipage}[b]{\textwidth}
%     \begin{center}
%       \includesvg[width=0.6\textwidth]{book1_proposition_47_figure_cover}
%     \end{center}
%   \end{minipage}
% \end{center}

% \vspace*{\fill}

% \begin{center}
%   \LARGE{Newell Jensen}\\
%   \LARGE{2023}
% \end{center}
% \newpage

% First Page
\pagecolor{background}
\thispagestyle{empty}

\hfill

\hfill

\hfill

\begin{flushright}
  \LARGE{Byrne's}\\
  \Huge{\color{cred}{E}\color{cblue}{u}\color{cyellow}{c}\color{cred}{l}\color{cblue}{i}\color{cyellow}{d}}
\end{flushright}

\vspace*{\fill}
\newpage

% Copyright and License
\thispagestyle{empty}

Newell Jensen\\
\url{https://newell.github.io}
% \the\textwidth
\vspace*{\fill}

\large{\textcopyright\ 2023 by Newell Jensen}

\hfill

\hfill

\centerline{\ccbyncsa}

\begin{center}
  \large{This work is licensed under the Creative Commons Attribution-NonCommercial-ShareAlike 4.0 International License (\href{https://creativecommons.org/licenses/by-nc-sa/4.0/}{CC BY-NC-SA 4.0}).}
\end{center}
\vspace*{\fill}
\newpage

% Dedication
\thispagestyle{empty}

\vspace*{\fill}
\begin{quote}
  \centering
  \Large{Dedicated to the \textsc{COL{\color{cred}{O}}{\color{cblue}{U}}{\color{cyellow}{R}}FUL} memory of my father, \LARGE{\textsc{William Henry Jensen}}.}\\
\end{quote}
\vspace*{\fill}
\newpage

% Preface
\pagenumbering{gobble}
\centerline{\LARGE{Preface}}

\hfill

\normalsize % Change back the size of the font


\hspace{2em}\textit{Let no one destitute of geometry enter my doors.}

\hspace{16em} --- Plato

\hfill

\begin{center}
  \textit{Sire, there is no royal road to geometry.}

  \hspace{10.5em}--- Euclid
\end{center}

\hfill

\begin{flushright}
  \textit{The mathematician's patterns, like the painter's or the poet's must be beautiful;}
  \textit{the ideas like the colours or the words, must fit together in a harmonious way.}
  \textit{Beauty is the first test: there is no permanent place in the world}\\
  \textit{for ugly mathematics.}
\end{flushright}

\hfill --- G. H. Hardy\\

Legend has it that the words, \textit{"Let no one destitute of geometry enter my doors"} were inscribed in the doorway entering Plato's Academy in Athens.  Euclid, an ancient Greek mathematician of Alexandria and a contemporary of Plato, considered the '\textit{Father of geometry}', is chiefly known for his treatise of 13 books called the \textit{Elements}.  A mathematical and logical masterpiece, the \textit{Elements} is a collection of definitions, postulates, propositions (consisting of theorems, problems and constructions), as well as the logical proofs of these propositions.  The \textit{Elements} has been referred to as the most successful and influential textbook ever written.

According to the ancient Greek historian Proclus, when the King of Egypt asked Euclid if there was an easier way to learn geometry, Euclid famously replied: \textit{"There is no royal road to geometry."} A well-known quote, often used to emphasize the importance of hard work, discipline, and persistence in the pursuit of knowledge and understanding, it suggests that there are no shortcuts to true understanding, and that the only way to master a subject is through diligent study and practice.

Enter Oliver Byrne.  An Irish-born civil engineer and surveyor, Byrne (1810-1880) is best known for his
illustrated edition of the first six books of Euclid’s \textit{Elements}, published in 1847 under the title
\textit{Byrne’s Euclid}. Byrne’s edition is noted for its distinctive use of colour-coded diagrams and symbols, intended to make the complex concepts of Euclidean geometry more accessible and understandable to a wider audience.  Each of the book’s geometric figures is rendered in bold primary colors, with different colors being used to distinguish between various parts of the figure and to indicate different types of lines and angles.  Byrne's edition of the \textit{Elements} was a commercial success and went through several editions in the years following its initial publication. While the book's approach to illustrating geometry was unconventional for its time, it has since become a popular and influential work in the field of graphic design, as well as a fascinating example of the intersection between mathematics, art, and visual communication.

Overall, \textit{Byrne's Euclid} represents a unique and innovative approach to the study of geometry -- one that combines technical precision with a bold and imaginative visual style. Inspiring generations of students and scholars to approach the study of mathematics with a sense of creativity and wonder, it
remains an important and enduring work in the history of mathematical literature.  The book's use of colorful diagrams and illustrations, along with concise and straightforward explanations, can make it easier for students to understand abstract concepts and develop a deeper appreciation for the beauty and elegance of mathematics. Additionally, studying \textit{Byrne's Euclid} can help students develop problem-solving skills and logical reasoning, which are valuable not just in mathematics but in many areas of life.

My first experience with \textit{Byrne’s Euclid} occurred while I was working through Euclid’s \textit{Elements} and searching on the Internet for more information about a particular proposition, when I came across Nicholas Rougeux's exquisite reproduction of Oliver Byrne's celebrated work --- \url{https://c82.net/euclid}.  Awed by the stunning beauty of the \textit{Elements} and its logical precision, as well as Byrne's masterful and imaginative approach, I was filled with inspiration to create this book.  Both G. H. Hardy's quote and \textit{Byrne's Euclid}, underscore the creative and aesthetic dimensions of mathematics. Hardy's quote highlights how mathematicians, like painters or poets, create enduring patterns with ideas, while \textit{Byrne's Euclid}, visually showcases the beauty of mathematical concepts through intricate illustrations. Together, they remind us that mathematics is not solely a logical pursuit, but also a richly imaginative and expressive one.  It is my hope that this rendition of \textit{Byrne's Euclid} continues in this spirit.

While Byrne's original work featured the elegant \textit{Caslon} typeface, I have chosen to use the open source \href{https://github.com/georgd/EB-Garamond}{\textit{EB Garmond}} typeface for my edition. While the two typefaces share many similarities in ligatures and glyphs, those who aren't typography experts may not even notice the difference.  For example, both \textit{Caslon} and \textit{EB Garamond} are serif typefaces, which means they have small decorative lines at the ends of each letter stroke. Classic and elegant, both have had a long history in printing and publishing.

In terms of their specific design features, these typefaces have similar letter shapes and proportions. For example, they both have a lowercase "a" with a curved tail, and a lowercase "g" with a descending loop. They also both have a tall and narrow uppercase "H", and a diagonal crossbar on the uppercase "A".  Finally, both typefaces feature ligatures (two or more letters that are joined together into a single glyph), such as "fi" and "fl", which have a similar design and placement in the two typefaces.

First time readers of \textit{Byrne's Euclid} should be made aware that the long s ({\color{cred}{ſ}}) is used throughout the book. The long s ({\color{cred}{ſ}}) is a letterform of the Latin alphabet that was commonly used in Europe from the Middle Ages until the 19th century. It looks like a lowercase "s", but with a longer, more elongated shape, resembling an "f" without its crossbar. In printed materials from the time, the long s was used in place of a normal "s" at the beginning or in the middle of a word, but not at the end of a word or after certain letters like "m", "n", or "u."

A stylistic convention, the use of the long s was thought to make text easier to read and more aesthetically pleasing, as it allowed letters to be more closely spaced and made words look more uniform in appearance. However, as printing technology evolved and more uniform letter spacing became possible, the long s fell out of use and was gradually replaced by the modern short "s" in the 19th century. Despite its decline in usage, the long s can still be found in some historic texts and remains a fascinating example of the evolution of written language over time.  Here is an illustration of the long s,
\begin{center}
  {\color{cred}{ſ}}orts $=$ sorts $\neq$ forts\\
  \textit{ca{\color{cred}{ſ}}e} $=$ \textit{case} $\neq$ \textit{cafe}
\end{center}

I would like to thank Nicholas Rougeux for giving me permission to use the scalar vector graphic images from his website in producing this book.  I would also like to thank all the logicians and mathematicians, past and present, whose shoulders we all stand on.  And to you, the reader.  It's my sincere hope and desire that you enjoy this book and the logical truths found herein.  I encourage you to study and practice the propositions so that you may walk your {\color{cred}{o}}{\color{cblue}{w}}{\color{cyellow}{n}} road to geometry.

\hfill

\begin{flushright}
  Newell Jensen, 2023
\end{flushright}

\newpage

\vspace*{\fill}

\begin{center}
  \textit{Reason is immortal, all else is mortal.}

  \hspace{8em}--- Pythagoras
\end{center}

\vspace*{\fill}

\newpage

% TOC
\tableofcontents

\newpage

% Introduction
\pagenumbering{roman}
\include{intro}
\pagenumbering{arabic}

% Book sections
\include{book1/book1}
\include{book2/book2}
\include{book3/book3}
\include{book4/book4}
\include{book5/book5}
\include{book6/book6}

\newpage

% The End
\thispagestyle{empty}
\vspace*{\fill}
\begin{center}
  \scshape{\LARGE{THE END.}}
\end{center}
\vspace*{\fill}

\end{document}